### 2023-03-29 main/chapter_experiment.tex
add met resolution

Since the $\twoljets$ analysis is most concerned with muons, electrons and
jets.
\TODO{and missing transverse momentum},
this section reviews \atlas's experimental resolution for these objects.

@@

\paragraph{Jets.}
Although charged hadrons are tracked in the inner detector, the total energy of
jets also depends vitally on neutral hadrons, which are only observed through
their hadronic interactions in the calorimeter system.
With less influence from tracking, calorimetry noise is the dominant
contribution to the experimental resolution on hadronic jets.
However, other noise contributions arising from electrical signals and pile-up
scale in proportion in $1/\pt$.
one can therefore again approximate \atlas's relative jet energy resolution as
\begin{equation}
\label{eqn:experiment_jet_resolution}
\frac{\sigma(\pt)}{\pt}
\approx
\frac{b_1}{\sqrt{\pt}}
\oplus
\frac{b_2}{\pt}
\oplus
b_3
,
\end{equation}
the same functional form as for electrons in
Equation~\ref{eqn:experiment_electron_resolution},
with
$b_1 \approx 1.0\,\sqrt{\eV[G]}$
$b_2 \approx 0.5\,\eV[G]$, and
$b_3 \approx 0.03$~\cite{atlas2021jetresolution}.
This formula approximately describes the measured jet resolutions of
$\frac{\sigma(\pt)}{\pt} \approx 20\%$ at $\pt = 30\,\eV[G]$,
$\frac{\sigma(\pt)}{\pt} \approx 8\%$ at $\pt = 200\,\eV[G]$, and
$\frac{\sigma(\pt)}{\pt} \approx 4\%$ at
$\pt = 2\,\eV[T]$~\cite{atlas2021jetresolution}.

@@

\paragraph{Missing transverse momentum.}
\TODO{}
% Reference from Holly
%  ATLAS collaboration. ET
% miss performance in the ATLAS detector using 2015–2016 LHC
% pp collisions. ATLAS-CONF-2018-023, 2016.

>>>>

Since the $\twoljets$ analysis is most concerned with muons, electrons and
jets and missing transverse momentum,
this section reviews \atlas's experimental resolutions for these objects.

@@

\paragraph{Jets.}
Although charged hadrons are tracked in the inner detector, the total energy of
jets also depends vitally on neutral hadrons, which are only observed through
their hadronic interactions in the calorimeter system.
With less influence from tracking, calorimetry noise is the dominant
contribution to the experimental resolution on hadronic jets.
However, other noise contributions arising from electrical signals and pile-up
scale in proportion in $1/\pt$.
one can therefore again approximate \atlas's relative jet energy resolution as
\begin{equation}
\label{eqn:experiment_jet_resolution}
\frac{\sigma(\pt)}{\pt}
\approx
\frac{b_1}{\sqrt{\pt}}
\oplus
\frac{b_2}{\pt}
\oplus
b_3
,
\end{equation}
the same functional form as for electrons in
Equation~\ref{eqn:experiment_electron_resolution},
with
$b_1 \approx 1.0\,\sqrt{\eV[G]}$
$b_2 \approx 0.5\,\eV[G]$, and
$b_3 \approx 0.03$.
With these parameters, formula approximately describes the measured relative
jet resolutions of
$\frac{\sigma(\pt)}{\pt} \approx 20\%$ at $\pt = 30\,\eV[G]$,
$\frac{\sigma(\pt)}{\pt} \approx 8\%$ at $\pt = 200\,\eV[G]$, and
$\frac{\sigma(\pt)}{\pt} \approx 4\%$ at
$\pt = 2\,\eV[T]$~\cite{atlas2021jetresolution}.

@@

\paragraph{Missing transverse momentum.}
Measurement resolutions of all other objects propagate into the missing
transverse momentum; this is necessary as it is not directly observed itself,
but derived from other objects' transverse momenta.
But collisions also emit significant amounts of momentum in softer particles,
which can leave tracks and calorimeter hits without being clustered into hard
physical objects such as jets or electrons.
The $\met$ resolution therefore depends not only on the resolutions of hard
objects, but also on the measurement of this soft contribution.
Various different $\met$ definitions, known as `working points', use different
quality criteria to define the $\met$ composition, and each working point
balances resolution against accuracy for different context-specific
objectives~\cite{ATLAS-CONF-2018-023, Pacey:2747774}.

Pile-up is a major contributor to the $\met$ resolution, because secondary
interactions produce both tracks and calorimeter clusters that might can be
indistinguishable from products of the hard-scatter primary vertex
(the main interaction of interest in a bunch crossing).
All misattributed momentum from these additional vertices causes $\met$ noise.
The \atlas\ $\met$ resolution is reported in two orthogonal transverse
components: $\sigma(\metx)$ and $\sigma(\mety)$.%
\footnote{%
The directional decomposition of $\ptmiss$ uncertainty is important for the
$\met$ significance variable $\metsig$, which we shall discuss in
Section~\ref{sec:2ljets_metsig}.
}%
\footnote{%
Despite their $E$ symbols, these variables $\metx$ and $\mety$ are momenta
because they have spatial directions.%
}
In reference samples of simulated $Z \to \mu\mu$ events, both transverse
$\met$ resolution components increase with the number of primary vertices in a
bunch crossing ($n_{PV}$), from
$\sigma(\metx) = \sigma(\mety) \approx 10\,\eV[G]$ at $n_{PV} = 1$, to
$\sigma(\metx) = \sigma(\mety) \approx 20\,\eV[G]$ at $n_{PV} = 30$.
Note that these resolutions vary by around $\pm 20\%$ with the $\met$
working point choice~\cite{ATLAS-CONF-2018-023}.
